---
title: "ESM 244 Lab Week 4 Key"
author: "Allison Horst"
date: "January 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab Week 4:

- Nonlinear least squares
- Panel regression level 0 example
- Shiny example with navbarPage

```{r}
library(tidyverse)
library(Ecdat)
library(plm)
library(lmtest)
library(car) #for plotting
```

###Part 1. Nonlinear least squares (logistic growth fitting)

a. Load dataset CellGrowth.csv. Create a scatterplot.

```{r get_data}

CellGrowth <- read_csv("CellGrowth.csv")

```

b. Look at it

```{r cell_graph} 

ggplot(CellGrowth, aes(x = Time, y = CellCount)) + geom_point()

```

Recall logistic growth equation: 

N(t) = A/(1+Be^-rt)

# Estimate for K/A: ~ 3700
# Estimate for N0: ~ 2000
# Estimate for B: ~ 0.85

BEstimate <- (3700 - 2000)/2000 # BEstimate = 0.85

c. Create a subset of the data that you think is just in the 'exponential growth phase.' Take the natural log of the count and create a basic scatterplot of time v ln(counts). 

```{r}

Graph2 <- plot(CellGrowth$Time[0:5],log(CellGrowth$CellCount[0:5])) # Create a basic scatterplot of time versus ln(CellCounts)

```

Then find the slope of that line to get 'r' (growth rate constant)

```{r}

r_est <- lm(log(CellGrowth$CellCount[1:5]) ~ CellGrowth$Time[1:5]) # R ~ 0.1035

```


d. Nonlinear least squares 

e. Model fitting 

```{r}

cell_fit <- nls(CellCount ~ A/(1+B*exp(-r*Time)), 
                start = list(A = 3700, B = 0.85, r = 0.1035), 
                data = CellGrowth, trace = TRUE) 

```


Uses the estimates as a start to find the best estimated values (by convergence) for A, B and r

8 iterations for convergence
Estimates, St. Errors and p-values

Parameters:
Estimate Std. Error t value Pr(>|t|)    
A 3.806e+03  5.029e+01   75.67  < 2e-16 ***
B 9.233e-01  5.887e-02   15.68 2.33e-09 ***
r 3.443e-01  3.219e-02   10.70 1.73e-07 ***

Can also find confidence intervals by confint(ModelFit)

```{r}

cell_ci <- confint(cell_fit)

```

f. Create new variables for A, B and r

```{r}

A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]

```

g. Create a new sequence containing a series of times over which you'll predict the cell count

```{r}

time_seq <- seq(0,20, length = 100)

```

h. Using the parameters (A, B, and r) and the time sequence, predict the cell counts for the logistic growth model

```{r}

cell_pred <- A/(1+B*exp(-r*time_seq)) # Logistic growth model

```

i. Bind together the time sequence data and the predictions data into a new data frame

```{r}

pred_df <- data.frame(time_seq, cell_pred)

```


j. Create a single graph in which you show the original data (as scatterplot points) and the predicted data (as a line graph)

```{r}

ggplot(CellGrowth, aes(x = Time, y = CellCount)) + 
  geom_point(colour = "blue", size = 3) + 
  theme_bw() +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred), colour = "orange", size = 1) + 
  xlab("Time (h)") +
  ylab("Cell Count") +
  ggtitle("Bacterial Growth")

```

###Part 2. Panel Regression - First Shot

The Data:

This data and example came from the Ecdat package in R: <https://cran.r-project.org/web/packages/Ecdat/Ecdat.pdf>

All you have to do is install and load the "Ecdat" package. The Cigarette dataset contains information on the number of cigarette packs per person across all 50 United States from 1985-1995.  We will look at the following columns:

- state (our "entity")
- year (time)
- packpc (the number of packs per capita- our DV)
- avgprs (average price during fiscal year, including sales taxes- our IV)

```{r data_setup, message=FALSE}

cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs) #only retaining the 4 columns we're interested in
  
```

### Entity Fixed Effects Model

Our research question is: How did the average price of cigarettes relate to the number of packs per person from 1985-1995 ON AVERAGE across all 50 states?

```{r cigs_model}

cigs_model <- plm(packpc ~ avgprs,
                data = cigs_panel,
                index = c("state","year"),
                model = "within")

cigs_model #So the number of packs per person decreased by an average of 0.364 when average price increased by 1 dollar across all 50 states

coeftest(cigs_model, vcov. = vcovHC(cigs_model, type = "HC1")) # coefficient is extremely significant (using robust SE)


```


But we know the number of packs per person has been going down across the United States for health reasons too, not just financial.  So we are going to use a time and entity fixed effects model.

### Time and Entity Fixed Effects Model


```{r et_cigs_model}

et_cigs_model <- plm(packpc ~ avgprs,
                data = cigs_panel,
                index = c("state","year"),
                model = "within",
                effect = "twoways")

et_cigs_model

coeftest(et_cigs_model, vcov. = vcovHC(et_cigs_model, type = "HC1")) #coefficient is still highly significant, but is a different value
```

### Which model is better?

Should we use time fixed effects, or just entity fixed effects?

```{r time_fe_test}

pFtest(et_cigs_model, cigs_model)

```

**Conclusion:** Yes, we should include time fixed effects.  This makes sense, because we know that cigarette use is decreasing universally over time due to so many factors - health, societal pressure, etc.  We would expect time to substantially influence the number of packs per person.


### Least Squares Dummy Variable Regression

So we've discovered that yes, the number of packs per person decreases with price on average across the US.  But we're interested in knowing more about the state-by-state variation- are there some states were price increase changes habits more than others?  How does the number of packs per capita vary by state?

We choose to look at the 5 states with the highest populations in the US from 1985-1995: California, Texas, New York, Florida, and Pennsylvania.

We can then look at the coefficients of each state to see how the packs per capita vary within states. Also, when we use a time + entity fixed effects model we can also see how packs per capita vary within years (logically, we would expect a general trend of decreasing cigarette use over time).

```{r data_setup2, message=FALSE}

cigs_panel2 <- Cigarette %>% 
                  select(state, year, packpc, avgprs) %>% 
                  #only retaining the 4 columns we're interested in
                  filter(state == "CA" | state == "TX"| state == "NY"| state == "FL"| state == "PA")
                  #only retaining the 5 states we're interested in
                  

```

```{r lsdv_cigs}

# Entity Fixed Effects Only

lsdv_cigs <- lm(packpc ~ avgprs + state, data = cigs_panel2) # Coefficient matches panel w/ model = "within". You can try this with the cigs_panel dataset if you want to double check the result

lsdv_cigs #Shows us that every state has a higher number of packs per person than California 

coeftest(lsdv_cigs, vcov. = vcovHC(lsdv_cigs, type = "HC1"))

# Time and Entity Fixed Effects - we should use this given our pFtest() results from earlier

te_lsdv_cigs <- lm(packpc ~ avgprs + state + factor(year), data = cigs_panel2) # Coefficient matches panel w/ model = "within" and effects = "twoways". You can try this with the cigs_panel dataset if you want to double check the result

te_lsdv_cigs #Similar result to entity fixed effects model. We also see sensible coefficients for each year - each year has a lower number of packs per person relative to 1985, as expected.

coeftest(te_lsdv_cigs, vcov. = vcovHC(te_lsdv_cigs, type = "HC1")) 

```

### Visualizing our data

Below is one way to create a plot of average price vs. packs per capita, separated by state.

First we need to add one extra argument to the dataset

```{r plot, warning=FALSE, message=FALSE}

cigs_panel_plot <- cigs_panel2 %>% 
                      droplevels() # need to include this for plotting purposes- otherwise R will continue                                      to try to plot all 50 levels of the state factor, even though                                             we've only selected 5 values 

levels(cigs_panel_plot$state) # check to make sure there are only 5 levels

scatterplot(cigs_panel_plot$packpc ~ cigs_panel_plot$avgprs|cigs_panel_plot$state, 
                boxplots=FALSE, 
                xlab="Average Price", 
                ylab="Packs Per Capita",
                smooth=FALSE, 
                legend.title = "State",
                ylim = c(55, 130)) 


# If you really wanted to do it for every state, you could...

scatterplot(cigs_panel$packpc ~ cigs_panel$avgprs|cigs_panel$state, 
                boxplots=FALSE, 
                xlab="Average Price", 
                ylab="Packs Per Capita",
                smooth=FALSE, 
                legend.title = "State",
                legend.coords = "right",
                legend.columns = 3,
                ylim = c(40, 200),
                xlim = c(80, 310)) 

```
